---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---


```{r, warning=False, message=False}
library(ggplot2)
library(GGally)
require(nnet)
library(randomForest)
library(lubridate)
require(MASS)
library(forcats)
library(pscl)
library(gridExtra)
library(ROCR)

US_Accidents <- read.csv("~/Downloads/Accidents Data CSV.csv",
                             stringsAsFactors=FALSE, header=T)
```


```{r}
summary(US_Accidents)
```

making data the right type. converting to Na and removing Na. 
new feature creature- ZipcodeLevels, timediffmin from existing features
```{r}
# Merge severity into two factors
US_Accidents$Severity <- factor(US_Accidents$Severity)
US_Accidents$Severity <- fct_collapse(US_Accidents$Severity, Low = c("1","2"), High = c("3","4"))


#Convert side to factor and remove empty
US_Accidents$Side <- factor(US_Accidents$Side)
levels(US_Accidents$Side)[1] <- NA
levels(US_Accidents$Side) <- c("Left", "Right")

US_Accidents$Amenity <- factor(US_Accidents$Amenity)
US_Accidents$Bump <- factor(US_Accidents$Bump)
US_Accidents$Crossing <- factor(US_Accidents$Crossing)
US_Accidents$Give_Way <- factor(US_Accidents$Give_Way)
US_Accidents$Junction <- factor(US_Accidents$Junction)
US_Accidents$No_Exit <- factor(US_Accidents$No_Exit)
US_Accidents$Railway <- factor(US_Accidents$Railway)
US_Accidents$Roundabout <- factor(US_Accidents$Roundabout)
US_Accidents$Station <- factor(US_Accidents$Station)
US_Accidents$Stop <- factor(US_Accidents$Stop)
US_Accidents$Traffic_Calming <- factor(US_Accidents$Traffic_Calming)
US_Accidents$Traffic_Signal <- factor(US_Accidents$Traffic_Signal)
US_Accidents$State <- factor(US_Accidents$State)
US_Accidents$Time.Interval <- factor(US_Accidents$Time.Interval)

US_Accidents$Timezone <- factor(US_Accidents$Timezone)
levels(US_Accidents$Timezone)[1] <- NA

# Convert ZipCode to 10 level factor
US_Accidents <- US_Accidents[!is.na(US_Accidents$Zipcode), ] # Remove rows without zipcode
US_Accidents$ZipcodeLevels[nchar(US_Accidents$Zipcode)==4] <-  0 # Convert all 4 digits zipcode to level 0
US_Accidents$ZipcodeLevels <- substr(US_Accidents$Zipcode, 1, 1) 
US_Accidents$ZipcodeLevels <- factor(US_Accidents$ZipcodeLevels) 

levels(US_Accidents$ZipcodeLevels)[1] <- NA # Convert empty zipcodelevel to NA
US_Accidents <- US_Accidents[!is.na(US_Accidents$ZipcodeLevels), ] # Remove rows with empty zipcodelevels

# Trim the characters after 5 digits -> if not will cause NA issues
US_Accidents$Zipcode <- substr(US_Accidents$Zipcode, 1, 5) 
US_Accidents$Zipcode <- as.numeric(US_Accidents$Zipcode) 


#Time difference between start and end time
US_Accidents$Start_Time <- mdy_hm(US_Accidents$Start_Time)
US_Accidents$End_Time <-  mdy_hm(US_Accidents$End_Time)
US_Accidents$timediff.min <- as.numeric(with(US_Accidents, difftime(End_Time,Start_Time,units="mins")))

sapply(US_Accidents, function(x) sum(is.na(x)))

US_Accidents = US_Accidents[-c(10,11)]
US_Accidents <- na.omit(US_Accidents)

str(US_Accidents)
names(US_Accidents)
```

```{r}
# Random Forest to find Best features for severity 
rf <- randomForest(Severity ~ ., data = US_Accidents)
varImpPlot(rf)
```


```{r}
summary(US_Accidents)
```

```{r}
str(US_Accidents)
```

```{r}
sapply(US_Accidents, function(x) sum(is.na(x))) # no Na's
```


```{r}
ggplot(US_Accidents, aes(x= Start_Lng, y = Start_Lat, color= Severity)) + geom_point(alpha = I(3/4)) 
```

From plotting the longitudinal and latitudinal points, we created a map which shows a clear distinction of where accidents of both severity levels occur based on geographic location. From this plot, we can see that low severity accidents are more probable, and are found within states (locally). On the other hand, high severity accidents occur more in the highways and interstates (eg. I-95 NY->FL).

```{r}
# checking correlations
mytable <- xtabs(~Severity + Side , data= US_Accidents)
ftable(mytable) # print table   if side is left data is more likely to be low severity i.e p = 0.93 .   
7583/(7583 + 513) # Severity data has more obs in 'low' category , also seen in ggpairs 
```

```{r}
mytable <- xtabs(~Severity + State, data= US_Accidents)

ftable(mytable) 
# we find high accidents in states : Texas = 2199, Florida= 1733, California= 1426
```

writing out FINAL CLEAN DATA
```{r}
write.table(US_Accidents, 
            file = "group6_cleandata.csv", 
            sep = "\t",
            row.names = FALSE)  
```


